---
title: "Prediction Assignment Writeup"
author: "MisterT1000"
date: "8/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Initial Setup 
I'm loading the caret and random forest libraries and setting the working directory to begin.
```{r initial_setup, message = FALSE}
library(caret)
library(randomForest)
setwd("~/test-repo/PracticalMachineLearningProject/")
```

## Importing and cleaning data
I imported the data and changed the NA's to 0. The next step is removing the data elements that are not actually predictors, such as the name, window, and timestamp columns. I also made sure to convert data that came up as factors back to the numeric types they should be.
```{r data, message = FALSE}
dataset <- read.csv("pml-training.csv")
dataset[is.na(dataset)] <- 0
dataset <- dataset[,-c(1:7)]
dataset[,sapply(dataset, class)=="factor"][-34] <- sapply(dataset[,(sapply(dataset, class)=="factor")],as.numeric)[-34]
```

I'm using the X,Y, and Z component predictors because the others are generally calculated from these and, as a result, my hypothesis is that those would not be as signinficant.
```{r predictors}
dataset <- dataset[,grepl("_x$|_y$|_z$|classe",names(dataset))]
```

I split my training data into a new training and validation set.
```{r training}
training <- createDataPartition(y = dataset$classe, p = 0.7, list = FALSE)
train <- dataset[training, ]
validation <- dataset[-training, ]
```

I'm using the random forest method to build my model.
```{r model}
modRF <- randomForest(classe~.,data=train)
```

Find the out of sample error rate with the validation set.
```{r validate}
pred <- predict(modRF,validation)
result <- confusionMatrix(pred,validation$classe)
result
```

Apply the model to the test data set to make a prediction.
```{r test}
finalTest <- read.csv("pml-testing.csv")
finalTest[is.na(finalTest)] <- 0
finalPred <- predict(modRF,finalTest)
finalPred
```